---
title: "Helpful functions in R 'meta' package"
author: "Douglas Mesadri Gewehr, MD - Meta-Analysis Academy"
output: 
  html_document: 
    toc: true
    toc_depth: 6
---

------------------------------------------------------------------------

### Loading Packages

------------------------------------------------------------------------

```{r Loading Packages, message=FALSE, warning=FALSE}
library(meta)
library(readxl)
```

------------------------------------------------------------------------

### Importing the Dataset

------------------------------------------------------------------------

The entire dataset was converted into a dataframe format for easier handling and compatibility with the R example analysis. The dataframe structure allows seamless data manipulation and analysis using R’s built-in functions and packages.

```{r Dataset, include=FALSE}
ma<-list(acm=structure(list(Author = c("Almendro-Delia et al.", "ATLANTIC Trial", 
"Danchin et al.", "Load&Go Trial", "CIPAMI Trial", "Fabris et al.", 
"Postma et al.", "Dorler et al.", "Lupi et al.", "Redfors et al.", 
"Yudi et al.", "Alexander et al.", "Fefer et al.", "Lev et al."
), Year = c(2015, 2014, 2018, 2013, 2012, 2022, 2014, 2011, 2016, 
2019, 2018, 2008, 2009, 2008), event.e = c(188, 30, 9, 2, 1, 
17, 46, 55, 5, 1960, 25, 12, 12, 1), n.e = c(3973, 906, 360, 
56, 164, 237, 1345, 1635, 143, 37840, 892, 1029, 217, 165), event.c = c(116, 
19, 18, 0, 4, 17, 35, 110, 13, 528, 96, 55, 6, 2), n.c = c(2076, 
952, 360, 56, 171, 300, 1096, 1076, 143, 6964, 1915, 2756, 166, 
127), p2y12_type = c("Clopidogrel", "Ticagrelor", "Clopidogrel", 
"Clopidogrel", "Clopidogrel", "Clopidogrel", "Clopidogrel", "Clopidogrel", 
"Ticagrelor", "Clopidogrel", "Clopidogrel", "Ticagrelor", "Clopidogrel", 
"Clopidogrel"), overal_risk_of_bias = c("Serious-risk of Bias", 
"Low-risk of Bias", "Low-risk of Bias", "Moderate-risk of Bias", 
"Low-risk of Bias", "Moderate-risk of Bias", "Serious-risk of Bias", 
"Moderate-risk of Bias", "Serious-risk of Bias", "Moderate-risk of Bias", 
"Moderate-risk of Bias", "Serious-risk of Bias", "Serious-risk of Bias", 
"Serious-risk of Bias"), `pretreatment_timing` = c("Pre-hospital", 
"Pre-hospital", "Pre-hospital", "Pre-hospital", "Pre-hospital", 
"Pre-hospital", "Pre-hospital", "Pre-hospital", "Pre-hospital", 
"Pre-hospital", "Pre-hospital", "In-hospital", "In-hospital", 
"In-hospital"), `%_HTN` = c(30, 28.8, 25.9, 52.3, 49.8, 30.5, 
20.3, 40.1, 60.6, 40.5, 43.4, 61.4, 32.6, 70.1)), row.names = c(NA, 
-14L), class = "data.frame"))
```

```{r View Dataset, echo=FALSE}
ma$acm
```

------------------------------------------------------------------------

### Performing Subgroup Analyses

------------------------------------------------------------------------

Before running any meta-analysis function (e.g., `metabin()`, `metacont()`, `metaprop()`, `metagen()`, or others), it is essential to reorder the levels of the subgroup variable in your dataset. The order of these levels determines how the subgroups will appear in the forest plot.

To adjust the order in which the subgroup levels will appear in the forest plot, use the `factor()` function:

```{r, Adjusting the order of subgroups}
ma$acm$overal_risk_of_bias = factor(ma$acm$overal_risk_of_bias, 
                                     levels = c("Low-risk of Bias", "Moderate-risk of Bias", "Serious-risk of Bias"))

ma$acm$p2y12_type = factor(ma$acm$p2y12_type, 
                            levels = c("Clopidogrel", "Ticagrelor"))

ma$acm$pretreatment_timing = factor(ma$acm$pretreatment_timing, 
                                    levels = c("Pre-hospital", "In-hospital"))
```

------------------------------------------------------------------------

#### First Step = Performing the Meta-Analysis

------------------------------------------------------------------------

To perform a subgroup analysis, use the `subgroup` (replaces the old `byvar` argument) argument in the meta-analysis function. This argument specifies the column in your dataset containing the subgroup definitions. The results will then be grouped by these subgroups and displayed in a structured format in the forest plot:

```{r, running the meta-analysis of subgroups}
m.acm <- metabin(event.e, n.e, event.c, n.c,
                 data = ma$acm,
                 method = "MH",
                 method.tau = "DL",
                 sm = "RR",
                 studlab = Author,
                 MH.exact = T,
                 print.subgroup.name = F)

summary(m.acm)
```

------------------------------------------------------------------------

#### Second Step = Updating the Meta-Analysis Running the Subgroup Analyses

------------------------------------------------------------------------

```{r}

m.acm_ROB = update(m.acm, subgroup = overal_risk_of_bias, print.subgroup.name=F)
m.acm_P2Y12 = update(m.acm, subgroup = p2y12_type, print.subgroup.name=F)
m.acm_timing = update(m.acm, subgroup = pretreatment_timing, print.subgroup.name=F)

m.acm_ROB

```

------------------------------------------------------------------------

#### Third Step = Creating the Forest Plot

------------------------------------------------------------------------

To create a forest plot with subgroup analysis, use the `forest()` function and specify the subgroup-related arguments:

-   `subgroup = TRUE`
-   `col.subgroup = "black"`
-   `prediction.subgroup = TRUE`

```{r Creating the Forest Plot, echo=TRUE, fig.height=8, fig.width=12, dpi=600}
forest(m.acm_ROB,
       smlab = "All-cause Mortality",
       layout = "Revman5",
       sortvar = TE,
       lab.e = "Experimental", label.left = "Favors Experimental",
       lab.c = "Control", label.right = "Favors Control",
       ff.lr = "bold",
       leftcols = c("studlab", "Year","event.e", "n.e", "event.c", "n.c", "w.random", "effect", "ci" ),
       leftlabs = c("Studies", "Year", NA, NA, NA, NA, NA, NA, NA),
       text.random = "Random effects model",
       random = TRUE, 
       common = FALSE,
       test.overall.random = TRUE,
       rightcols = FALSE,
       colgap = "3mm",
       fs.heading = 12,
       fs.study = 12, 
       fs.hetstat = 10, 
       digits = 2,
       digits.pval = 2,
       pooled.events = TRUE,
       pooled.totals = TRUE,
       print.I2.ci = TRUE,
       col.square="darkcyan", col.square.lines="black",
       prediction = TRUE, col.predict = "#CEF2EE",col.predict.lines = "black", ff.predict = 1,
       subgroup = T,
       col.subgroup = "black",
       prediction.subgroup = TRUE)
```

------------------------------------------------------------------------

#### Fourth Step = Generating a Forest Plot With Results of Several Subgroup Analyses

------------------------------------------------------------------------

```{r, Updating the subgroup meta-analysis}
m.acm_ROB = update(m.acm, subgroup = overal_risk_of_bias, print.subgroup.name=T, subgroup.name = "Overall Risk-of-bias")
m.acm_P2Y12 = update(m.acm, subgroup = p2y12_type, print.subgroup.name=F, subgroup.name = "P2Y12 Inhibitors")
m.acm_timing = update(m.acm, subgroup = pretreatment_timing, sprint.subgroup.name=F, subgroup.name = "Pretreatment Timing")
```

```{r, Running the metabind function}

m.acm.subgroups= metabind(m.acm_ROB, m.acm_P2Y12, m.acm_timing, 
                          random = T,
                          common = F)

m.acm.subgroups
```

```{r, creating forestplot with several subgroup analyses, fig.height=5, fig.width=9, dpi=600}

forest(m.acm.subgroups,
       layout = "meta",
       smlab = "All-cause Mortality",
       lab.e = "Experimental", label.left = "Favors Experimental",
       lab.c = "Control", label.right = "Favors Control",
       ff.lr = "bold", col.subgroup = "black", col.square = "black",
       leftcols = c("studlab", "k", "effect", "ci"),
       rightcols = c("pval.Q.b"),
       overall.hetstat=FALSE, overall=F)       

```
